### 2013-07-19 abstract.tex
Initial commit.

% This is abstract.tex

The abstract should be a brief summary of your results that grabs the
reader's attention.  It is the first thing the reader will see, and you
must convince the reader that the rest of your paper is worthwhile.  In
the best of all possible worlds, \LaTeX\ will take care of all the
formatting for you.  Let $v = (v_0, v_1, \ldots, v_n)$.  If your
abstract contains nothing but words, no one would know that you're using
\LaTeX!