### 2013-07-19 appa.tex

% This is appa.tex
%
% This file is for appendices.  If you don't have any appendices, delete
% the percent sign from the beginning of the next line.
%\endinput

\appendix

% Add additional appendices by replicating the line below.

\section{Title of Appendix}

Appendices may appear after the paper proper.  Appendices may hold
extra information that would interrupt the flow of the paper and that
is not absolutely necessary for the reader to appreciate the work.
For example, a large number of related figures or a mathematical
derivation could go nicely in an appendix.