### 2013-07-19 summary.tex

% This is summary.tex

The non-technical summary is like an abstract, but without any technical
terminology. So instead of $v = (v_0, v_1, \ldots, v_n)$, you might say
something along the lines of ``pick a point in high-dimensional space.''
Don't worry about this section yet; you'll receive more guidance
later on, when you're near the end of your mentorship.