### 2013-07-24 appa.tex
Before meeting Ben.

% This is appa.tex
%
% This file is for appendices.  If you don't have any appendices, delete
% the percent sign from the beginning of the next line.
\endinput

\appendix

% Add additional appendices by replicating the line below.

\section{Casework for Theorem \ref{thm:count_maps}}
\label{appa:count_maps}
Compute like hell.