### 2013-07-25 summary.tex

% This is summary.tex

We consider maps between structures called Bott-Samelson bimodules, and study these maps using special diagrams.  We want to find a formula for the product of two such maps.  This paper gives a formula for the uniform case, and provides restrictions on the possible products for an ``alternating'' case.  We also count the number of maps between any two given Bott-Samelson bimodules.