### 2013-07-29 abstract.tex
Hell Week Monday 11AM update.

% This is abstract.tex

This paper considers left/right action-preserving morphisms between Bott-Samelson bimodules from a combinatorial perspective; specifically, using a diagrammatic representation.  We examine the maps formed by a basis known as Libedinsky's light leaves, which constructs a map based on a string $\underline r$, composed of the letters $s$ and $t$, and a binary string $\underline b$.  The eventual goal is to find a formula for the composition of two such maps based on their binary strings $\underline b$.  The paper provides a complete formula for the case $\underline r = sss\dots$, and shows that the possible structures of maps when $\underline r = ststs\dots$ are extremely limited.  We additionally provide a formula for the number of maps between two arbitrary Bott-Samelson bimodules.