### 2013-07-30 summary.tex

% This is summary.tex

We take a certain class of maps and represent them with colored diagrams.  These colored diagrams provide a combinatorial way for us to study and compute these maps.  Now an algorithm of Libedinsky allows us to construct such diagrams from a string of $1$'s and $0$'s; furthermore, we can ocmpose two such diagrams together to obtain a third diagram.   We develop tools to compute this third diagram based on the original binary strings.

%TODO write one